A Convolutional Neural Network for Outsole Recognition\tnoteref{t1}

### 2019-06-03 paper/paper.Rnw
structure/outline comments

In this paper, we discuss an alternate approach that uses an additional model layer to transfer the feature vector output into a vector of probabilities representing the detection of geometric elements in shoe tread images. Working within the feature space used by forensic examiners allows us to augment human-identified features with model output, assessing similarity of different shoe images on a feature set that is explicitly relevant to the domain.

\section{Materials and Methods}
\paragraph{Transfer Learning}
Neural networks are composed of sets of layers; the early layers, which are called the model base, contain feature detectors, and the final layers make up the classifier, or model head, which connects meaningful features to classification labels. A relatively simple convolutional neural network which processes image data may contain more than 14 million parameters in the model base and an additional 120 million parameters in the model head; to train a model of this nature requires millions of labeled images and a significant amount of computational power. Assembling data to train a network from scratch is often a nearly impossible task; the process of \emph{transfer learning} is a natural solution to this common problem. Transfer learning leverages the modularity of neural networks, that is, that the model base can be separated from the classifier which produces predictions.

@@

\paragraph{Annotated Training Data}

@@

\section{Results}

@@

\subsection{Model Accuracy}

@@

\section{Conclusions}

The goal of this research was to develop a method to automatically identify geometric class characteristics of shoe outsoles. Such a method takes the ``impossible" problem of quantifying outsole features in a given population and lays the groundwork for a tractable solution. A set of geometric class characteristics was defined to both broadly classify a large variety of shoes and to narrow down similarity into a manageable number and type of features for further use in a shoeprint analysis. The final set of features used in this research achieves both goals quite well, and has potential for much finer-grained analysis when recognized features are combined with spatial information and relationships.

>>>>

In this paper, we discuss an alternate approach that uses an additional model layer to transfer the feature vector output into a vector of probabilities representing the detection of geometric elements in shoe tread images. Working within the feature space used by forensic examiners allows us to augment human-identified features with model output, assessing similarity of different shoe images on a feature set that is explicitly relevant to the domain."
\svp{Add 1-2 sentences connecting feature sets to the goal of reducing workload - possibly cite solemate as an example of manual feature identification}

\section{Materials and Methods}
\svp{Short bit (1-2 sentences) about low-level feature detectors and why they fail with this problem.}

\svp{Motivate the use of CNNs directly - complex model, higher level of feature aggregation. Mention pre-trained networks to transition.}

\svp{VGG16/ImageNet specifically. Note that the structure of the model is distinct from the fitted weights.}

\svp{Transfer Learning - using CNNs with less data. Make the point that this is what makes our method different from some of the others we've already cited - training a new classifier allows us to work within already-existing semantic contexts that examiners use.}

\paragraph{Transfer Learning}
Neural networks are composed of sets of layers; the early layers, which are called the model base, contain feature detectors, and the final layers make up the classifier, or model head, which connects meaningful features to classification labels. A relatively simple convolutional neural network which processes image data may contain more than 14 million parameters in the model base and an additional 120 million parameters in the model head; to train a model of this nature requires millions of labeled images and a significant amount of computational power. Assembling data to train a network from scratch is often a nearly impossible task; the process of \emph{transfer learning} is a natural solution to this common problem. Transfer learning leverages the modularity of neural networks, that is, that the model base can be separated from the classifier which produces predictions.

@@

\svp{Moving towards ... how did we train the model}
\paragraph{Annotated Training Data}

@@

\svp{This section header might need to move? Haven't decided yet...}
\section{Results}

@@

\svp{Starting into How CoNNOR performs}
\subsection{Model Accuracy}

@@

\section{Conclusions}
\svp{Rework this to address 3 main conclusions - 1: NNs work for automatic feature ID, 2: Different levels of feature variability within classes, and 3: correcting for contrast is important.}
The goal of this research was to develop a method to automatically identify geometric class characteristics of shoe outsoles. Such a method takes the ``impossible" problem of quantifying outsole features in a given population and lays the groundwork for a tractable solution. A set of geometric class characteristics was defined to both broadly classify a large variety of shoes and to narrow down similarity into a manageable number and type of features for further use in a shoeprint analysis. The final set of features used in this research achieves both goals quite well, and has potential for much finer-grained analysis when recognized features are combined with spatial information and relationships.
